### 2018-07-08 latex/discussion.tex
Started work on discussion.
As suggested by Mathias Z., I started with a list of points to discuss (mostly his comments until now).

The dynamic interactions among metabolism, signal transduction, and gene regulation are still not very well understood. New approaches are required to disentangle the underlying network, helping us to understand how alterations affect human physiology. We here combined phylogenetic analysis and mathematical modelling of the NAD biosynthesis and consumption pathway to better understand the evolution of this pathway and especially the role of the methyltransferase NNMT. Our analyses reveal that NNMT plays a vital role for the diversification of NAD-dependent signalling reactions and potentially also for NAD homeostasis. We show that the presence of NNMT enables higher NAD consumption rates in comparison to NamPRT or NADA alone. It furthermore appears that NNMT makes both NAD concentration and NAD consumption relatively independent of other processes requiring NAD such as cell growth. We furthermore suggest that NNMT exerted an evolutionary pressure on NamPRT, driving the development of the high affinity of NamPRT for its substrates found in the human enzyme \cite{Burgos2008}. These findings shed new light on the role of NNMT, which has earlier been recognized as potential marker for some types of cancer (e.g. \cite{Okamura1998}). Increased NNMT expression in cancer might thus serve to remove Nam produced by increased NAD-dependent signalling. Further analysis is required to better understand which NAD-dependent signalling reactions are affected.

@@

We here neglected the potential effects of the methyl donor \textit{S}-adenosyl methionine (SAM) and its precursor methionine. Changing SAM availability is most likely contributing yet another regulatory level for Nam availability and thus an additional interaction point between gene regulation and metabolism. It has been shown earlier that NNMT expression influences protein methylation dependent on methionine availability \cite{Ulanovskaya2013}. One of the challenges in the analysis of this aspect is the availability of \textit{in vivo} concentration measurements for SAM and the large amount of reactions using it as substrate. A simplified mathematical model of this pathway has been published earlier \cite{Reed2004} providing the basis for future studies.

>>>>

\subsection{Points to discuss}

+ NNMT evolution.

--> Why was the yeast MT defined as NNMT?

--> Are there any ideas where the NNMT came from?

--> Any information on the methylation activity (also in comparison to others)? Does it fit to the found kinetic functions?

+ NADA disappears but rest of pathway stays. Why?

--> NADS needed for \textit{de~novo} pathway

--> NMNAT also needed in NamPRT and \textit{de~novo} pathway

-----> Is the \textit{de~novo} pathway still necessary? If we take up enough Vit B3 with the diet, why \textit{de~novo}?

--> NAPRT needed for NA that is taken up with the diet (and if NA is processed, the other two enzymes are also needed to continue processing)

+

\subsection{Old discussion}

The dynamic interactions among metabolism, signal transduction, and gene regulation are still not very well understood. New approaches are required to disentangle the underlying network, helping us to understand how alterations affect human physiology. We here combined phylogenetic analysis, mathematical modelling, and experimental evidence to better understand the evolution of the NAD biosynthesis and consumption pathways and especially the role of the methyltransferase NNMT. Our analyses reveal that NNMT plays a vital role for the diversification of NAD-dependent signalling reactions and potentially also for NAD homeostasis. We show that the combination of NNMT and NamPRT enables higher NAD consumption rates in comparison to NamPRT or NADA alone. It furthermore appears that NNMT makes both NAD concentration and NAD consumption relatively independent of other processes requiring NAD such as cell growth. In addition, we suggest that NNMT exerted an evolutionary pressure on NamPRT, driving the development of the high affinity of NamPRT for its substrates found in the human enzyme \cite{Burgos2008}. These findings shed new light on the role of NNMT, which has earlier been recognized as potential marker for some types of cancer (e.g. \cite{Okamura1998}). Increased NNMT expression in cancer might serve to remove Nam produced by increased NAD-dependent signalling. Further analysis is required to better understand which NAD-dependent signalling reactions are affected.

@@

We neglected the potential effects of the methyl donor \textit{S}-adenosyl methionine (SAM) and its precursor methionine in our models. Changing SAM availability is most likely contributing yet another regulatory level for Nam availability and thus an additional interaction point between gene regulation and metabolism. It has been shown earlier that NNMT expression influences protein methylation dependent on methionine availability \cite{Ulanovskaya2013}. One of the challenges in the analysis of this aspect is the availability of \textit{in~vivo} concentration measurements for SAM and the large amount of reactions using it as substrate. A simplified mathematical model of this pathway has been published earlier \cite{Reed2004} providing the basis for future studies.
